---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# GoldinGroundOz

<!-- badges: start --> <!-- badges: end -->
GoldinGroundOz is a small exploratory R package that demonstrates the long-term relationship between Australia's detached house prices, household disposable income, and average capital gain rate changes.
It provides two simulated datasets and illustrative examples that explore one of the hidden forces behind Australia’s surging housing prices — the shifting patterns of capital gain rates, or how “gold lies beneath the ground.”

## Installation

You can install the development version of GoldinGroundOz from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("ETC5523-2025/assignment-4-packages-and-shiny-apps-Acrea-QwQ")
```

## Example

A simple example of exploring the included datasets:

```{r example}
library(GoldinGroundOz)

# Load the datasets
data("housepr_income")
data("capital_gain")

# Quick overview
head(housepr_income)
head(capital_gain)

# Plot house prices vs disposable income
plot(
housepr_income$year, housepr_income$house_price,
type = "l", col = "gold", lwd = 2,
ylab = "Index (1989–90 = 100)", xlab = "Year"
)
lines(housepr_income$year, housepr_income$income, col = "pink", lwd = 2)
legend("topleft", legend = c("House Prices", "Disposable Income"),
col = c("gold", "pink"), lwd = 2)
```
